### 2013-03-02 trunk/eusflat2013/preliminaries.tex
updated eusflat

This section introduces the framework of the ill-known contraints. First of all, some preliminary concepts about possibility theory is explained. Next to that, the representation of time intervals by using the presented framework is done. Finally, the evaluation of the Allen's relationships between a crisp time interval and a possibilistic time interval is provided.

\subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
Informally speaking, a possibilistic variable is a variable that takes only one value but that, for some reason, this value is unknown. This is rather usual for historical events. For example, the conquest of Granada by the Christians' kings took place on February of 1492, but the exact day is not precissely known.
A more formal definition of possibilistic variables is given as follows.
%Possibilistic variables rely on possibility theory \cite{Dubois1988a}. A \emph{possibilistic variable} is defined as follows \cite{Pons2011}.

@@

For example, consider an ill-known value $X$ given by a triangular possibility distribution $\left[15, 1, 1\right]$. The ill-known contraint for all the points greater than $X$ is given by $C \triangleq (>, X)$. See Figure \ref{fig:ikc-greater}.

\begin{figure}[h]
   \centering
   \includegraphics[scale=1.5]{graphs/gt.eps}
   \caption{An ill-known value $X = \left[15, 1, 1\right]$ and an ill-known contraint $C \triangleq (>, X)$. }
   \label{fig:ikc-greater}
 \end{figure}

Several ill-known constraints can be aggregated by using the possibilistic extension of boolean operators: and ($\wedge$), or ($\vee$), not ($\neg$). The possibility and neccesity for these operators are computed as follows.

@@

 \begin{figure}[h]
   \centering
   \includegraphics[scale=1.5]{graphs/ill-known-ti.eps}
   \caption{An ill-known time interval $J = \left[X, Y  \right]$ given by two ill-known values. }
   \label{fig:ill-known-ti}
 \end{figure}

An ill-known time interval is defined by two ill-known values (see Figure \ref{fig:ill-known-ti}). Therefore, both starting and ending points are not precissely known. The following is the notation for an ill-known time interval.

>>>>

This section introduces the framework of the ill-known constraints. First of all, some preliminary concepts about possibility theory is explained. Next to that, the representation of time intervals by using the presented framework is done. Finally, the evaluation of the Allen's relationships between a crisp time interval and a possibilistic time interval is provided.

\subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
Informally speaking, a possibilistic variable is a variable that takes only one value but that, for some reason, this value is unknown. This is rather usual for historical events. For example, the conquest of Granada by the Christians' kings took place on February of 1492, but the exact day is not precisely known.
A more formal definition of possibilistic variables is given as follows.
%Possibilistic variables rely on possibility theory \cite{Dubois1988a}. A \emph{possibilistic variable} is defined as follows \cite{Pons2011}.

@@

For example, consider an ill-known value $X$ given by a triangular possibility distribution $\left[15, 1, 1\right]$. The ill-known constraint for all the points greater than $X$ is given by $C \triangleq (>, X)$. See Figure \ref{fig:ikc-greater}.

\begin{figure}[h]
   \centering
   \includegraphics[scale=1.5]{graphs/gt.eps}
   \caption{An ill-known value $X = \left[15, 1, 1\right]$ and an ill-known constraint $C \triangleq (>, X)$. }
   \label{fig:ikc-greater}
 \end{figure}

Several ill-known constraints can be aggregated by using the possibilistic extension of boolean operators: and ($\wedge$), or ($\vee$), not ($\neg$). The possibility and necessity for these operators are computed as follows.

@@

 \begin{figure}[h]
   \centering
   \includegraphics[scale=1.5]{graphs/ill-known-ti.eps}
   \caption{An ill-known time interval IKTI $J = \left[X, Y  \right]$ given by two ill-known values. }
   \label{fig:ill-known-ti}
 \end{figure}

\begin{definition}
An ill-known time interval \emph{(IKTI)} is defined by two ill-known values (see Figure \ref{fig:ill-known-ti}). Therefore, both starting and ending points are not precisely known. The following is the notation for an IKTI.

### 2013-03-02 trunk/eusflat2013/proposal.tex
updated eusflat

Two frameworks for dealing with uncertainty in time intervals have been presented. In this section, we will study the equivalences between an uncertain time interval (UTI) and an ill-known time interval (IKTI). Nevertheless, there are some diferences in the semantics of UTIs and IKTIs which are also explained. Next to that, the computation of the Allen's Relations are compared.

@@

Where $\mu_S, \mu_E$ are triangular membership functions. Notice that the main diference between an UTI and an IKTI is a matter of interpretation. IKTIs  keep both possibility distributions for both starting and ending poins, while UTIs take a t-norm (the minimum) of both possibility distributions (see Eq. \eqref{eq:possibility-uti}). That is, not all the crisp intervals between the boundaries of the starting and ending point are candidate intervals for the UTI.

@@

It is possible to obtain the same interpretation for an IKTI, but some ill-known constraints must be defined over the starting and ending ill-known values $X$, $Y$. In \cite{Pons2011}, two constraints are defined to determinte whether a crisp time interval $[a, b]$ resides between the boundaries of $J$. In this case, all the crisp intervals whose boundaries are within the boundaries of $X$ and $Y$ are considered to be a candidate interval. Nevertheless, if we want to obtain the same interpretation as in the UTI, the following two ill-known constraints must be defined:

@@

\begin{itemize}
 \item In TM, the Allen's Relations $R \in \left \lbrace B, O, C, D, OB, A \right \rbrace$ between an interval point $J$  and an UTI $\pi_I$ are computed by Eq. \eqref{eq:possibility-iz-uti-certain}. Therefore, for the Allen's Relations in $R$, 
 possibility that an Allen Relation between an interval point $J$  and an UTI $\pi_I$ is $1$ if the interval point $J$ is in the Allen Relation, or $0$ in other case.
\item In the TM, the other nine Allen's Relations between an interval point $J$ and an UTI $\pi_I$ are graded from $1$ to $0$ and computed by using the Eq. \eqref{eq:possibility-iz-uti}. 
\item In the ill-known constraint framework, all the thirdteen Allen's Relations are computed by using an aggregation of ill-known constraints (see Table \ref{tab:allen-relations}). Therefore, the possibility that a time point $J$ is in a given Allen Relation is always graded from $1$ to $0$.
\end{itemize}

@@

The following table stablishes the correspondences between the Allen's relations in the TM and in the ill-known constraints framework.

@@

>>>>

Two frameworks for dealing with uncertainty in time intervals have been presented. In this section, we will study the equivalences between an uncertain time interval (UTI) and an ill-known time interval (IKTI). Nevertheless, there are some differences in the semantics of UTIs and IKTIs which are also explained. Next to that, the computation of the Allen's Relations are compared.

@@

Where $\mu_S, \mu_E$ are triangular membership functions. Notice that the main difference between an UTI and an IKTI is a matter of interpretation. IKTIs  keep both possibility distributions for both starting and ending poins, while UTIs take a t-norm (the minimum) of both possibility distributions (see Eq. \eqref{eq:possibility-uti}). That is, not all the crisp intervals between the boundaries of the starting and ending point are candidate intervals for the UTI.

@@

It is possible to obtain the same interpretation for an IKTI, but some ill-known constraints must be defined over the starting and ending ill-known values $X$, $Y$. In \cite{Pons2011}, two constraints are defined to determinate whether a crisp time interval $[a, b]$ resides between the boundaries of $J$. In this case, all the crisp intervals whose boundaries are within the boundaries of $X$ and $Y$ are considered to be a candidate interval. Nevertheless, if we want to obtain the same interpretation as in the UTI, the following two ill-known constraints must be defined:

@@

\begin{itemize}
 \item In TM, the Allen's Relations $R \in \left \lbrace B, O, C, D, OB, A \right \rbrace$ between an interval point $J$  and an UTI $\pi_I$ are computed by Eq. \eqref{eq:possibility-iz-uti-certain}. Therefore, for the Allen's Relations in $R$, 
 possibility that an Allen Relation between an interval point $J$  and an UTI $\pi_I$ is $1$ if the interval point $J$ is in the Allen Relation, or $0$ in other case.
\item In the TM, the other nine Allen's Relations between an interval point $J$ and an UTI $\pi_I$ are graded from $1$ to $0$ and computed by using the Eq. \eqref{eq:possibility-iz-uti}. 
\item In the ill-known constraint framework, all the thirteen Allen's Relations are computed by using an aggregation of ill-known constraints (see Table \ref{tab:allen-relations}). Therefore, the possibility that a time point $J$ is in a given Allen Relation is always graded from $1$ to $0$.
\end{itemize}

@@

The following table establishes the correspondences between the Allen's relations in the TM and in the ill-known constraints framework.

@@

We will illustrate the calculation of one Allen Relation with an example.
\begin{example}
Consider a crisp time interval $J = [1, 3]$ the two ill-known time points $S, E$ that define a time interval. 
We want to obtain the possibility that $J$ is before the time interval.
The points are defined by two triangular membership functions $\mu_S$, $\mu_E$.

\begin{align}
\label{eq:ex-point}
\pi_S &= \mu_S = \left[2,5,7 \right] \\
\pi_E &= \mu_E = \left[9,12,15 \right]
\end{align}

Then, it is possible to define both an IKTI $I$ and an UTI $\pi_I$.

\begin{align}
\label{eq:ex-uti-ikti}
I = \left[S, E \right]\\
\mu_I = \left(\mu_S, \mu_E \right)
\end{align}

First we will obtain the possibility for the UTI $\pi_I$. The relation possibly meets has three candidates: Before, Meets and Overlaps.

\begin{align}
  \mu_{\left(J R \pi_I \right)}(B) &= \sup_{a < e} \pi_S(a)\\
\mu_{\left(J R \pi_I \right)}(M) &= \pi_S(e)\\
\mu_{\left(J R \pi_I \right)}(O) &= \sup_{a > e} \pi_S(a)
\end{align}

\begin{align}
\mu_{\left(J R \pi_I \right)}(B) &= 0.67\\
\mu_{\left(J R \pi_I \right)}(M) &= 0.33\\
\mu_{\left(J R \pi_I \right)}(O) &= 0
\end{align}

In order to compute the possibility  for the before relation between the crisp interval $J$ and the IKTI $I$, a ill-known constraint is needed.

\begin{equation}
 C_1\stackrel{\triangle}{=} \left(<,X\right)
\end{equation}

Then the possibility is computed as follows:

\begin{align}
\Pos \left(C_1 (J) \right) &= \sup_{a < e} \pi_S(a)\\
\Pos \left(C_1 (J) \right) &= 0.67 
\end{align}

\end{example}

### 2013-03-02 trunk/eusflat2013/triangularmodel.tex
updated eusflat

\subsection{\label{subsec:visualization-tm}Visualization within the triangular model}
Time intervals are usually represented in a one dimensional scale. This is due to its linear characteristic. This approach has also been used in Figure\ref{fig:allen-relationships} and \ref{fig:ill-known-ti}. The dificulty of accurate visual analisys on the temporal data arises when the number of time intervals increase.

@@

\begin{figure}[h]
   \centering
   \includegraphics[width=0.8\columnwidth]{graphs/allencrc.eps}
   \caption{CRZs corresponding to the thirdteen Allen's relations. The following shorthand notations are used: equals (E), starts (S), started by (SB), finishes (F), finished by (FB), meets (M), met by (MB), overlaps (O), overlapped by (OB), during (D), contains (C), before (B) and after (A).  }
   \label{fig:allencrz}
 \end{figure}

@@

\subsection{\label{subsec:relational-info-uti}Obtaining Relational Information}
In order to make temporal reasoning with UTIs, in the triangular model, the thirteen Allen's Relations, and its respectives Crisp Relational Zones (CRZ) have been generalized. In this case, we obtain fifteen Uncertain Relational Zones (URZ) , shown in Fig. \ref{fig:urz} and in Table \ref{tab:urz}.

@@

\begin{enumerate}
 \item Draw the lines $L_s, L_e$ that represent the interval point $J$, with the procedure shown un Section \ref{subsec:visualization-tm}. 
\item Draw the orthogonal lines from $L_s, L_e$. 
\item One or several of these lines could intersect with the UIZ of $\pi_I$. Then, the UIZ is divided into as many different subzones $\pi_I^{R'}$ as possible candidate Allen relations $R'$ between $J$ and $\pi_I$ exist.
\item $\pi_I^{R'}$ is the zone that correspond to a candidate Allen Relation $R'$.
\end{enumerate}
The possibility that a candidate relation $R'$ is the actual relation between $J$ and $\pi_I$ is given by:
\begin{equation}
 \label{eq:possibility-iz-uti}
\mu_{\left(J R \pi_I \right)} \left(R' \right) = \sup_{I \in \pi_I^{R'}} \pi_I(I).
\end{equation}
\end{itemize}

>>>>

\subsection{\label{subsec:visualization-tm}Visualization within the triangular model}
Time intervals are usually represented in a one dimensional scale. This is due to its linear characteristic. This approach has also been used in Figure\ref{fig:allen-relationships} and \ref{fig:ill-known-ti}. The difficulty of accurate visual analysis on the temporal data arises when the number of time intervals increase.

@@

\begin{figure}[h]
   \centering
   \includegraphics[width=0.8\columnwidth]{graphs/allencrc.eps}
   \caption{CRZs corresponding to the thirteen Allen's relations. The following shorthand notations are used: equals (E), starts (S), started by (SB), finishes (F), finished by (FB), meets (M), met by (MB), overlaps (O), overlapped by (OB), during (D), contains (C), before (B) and after (A).  }
   \label{fig:allencrz}
 \end{figure}

@@

\subsection{\label{subsec:relational-info-uti}Obtaining Relational Information}
In order to make temporal reasoning with UTIs, in the triangular model, the thirteen Allen's Relations, and its respective Crisp Relational Zones (CRZ) have been generalized. In this case, we obtain fifteen Uncertain Relational Zones (URZ) , shown in Fig. \ref{fig:urz} and in Table \ref{tab:urz}.

@@

\begin{enumerate}
 \item Draw the lines $L_s, L_e$ that represent the interval point $J$, with the procedure shown in Section \ref{subsec:visualization-tm}. 
\item Draw the orthogonal lines from $L_s, L_e$. 
\item One or several of these lines could intersect with the UIZ of $\pi_I$. Then, the UIZ is divided into as many different subzones $\pi_I^{R'}$ as possible candidate Allen relations $R'$ between $J$ and $\pi_I$ exist.
\item $\pi_I^{R'}$ is the zone that correspond to a candidate Allen Relation $R'$.
\end{enumerate}
The possibility that a candidate relation $R'$ is the actual relation between $J$ and $\pi_I$ is given by:
\begin{equation}
 \label{eq:possibility-iz-uti}
\mu_{\left(J R \pi_I \right)} \left(R' \right) = \sup_{I \in \pi_I^{R'}} \pi_I(I).
\end{equation}
\end{itemize}
